An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2017-07-02 uai2017/paper.tex
Updated UAI paper with reviewer comments, +copyright form

We note that this approach is fundamentally different from prior
work. It makes no assumptions about the form of, and requires no global bounds on the
posterior parameter distribution. It is exact in the
limit as batch size increases by the Central Limit Theorem. This is
not true of ~\citep{cutting_mh_2014} and~\citep{icml2014c1_bardenet14}
which use tail bounds and provide only approximate tests even with
arbitratily large batches of data. Our test is also exact under the
assumptions of \citep{cutting_mh_2014} that the log probability ratios
of batches are normally distributed about their mean.  Rather
than tail bounds, our approach uses moment estimates from the data to
determine how far the minibatch posteriors deviate from a normal
distribution.  These bounds carry through to the overall accuracy of
the test.

@@

\subsection{BARKER (LOGISTIC) ACCEPTANCE FUNCTION}\label{ssec:barker_function}
For our new MH test we use the Barker logistic~\citep{Barker65}
function: $g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows
that it satisfies the condition in Lemma~\ref{lem:detailed_balance}.
It is slightly less efficient than the Metropolis test, since its
acceptance rate for vanishing likelihood difference is 0.5. However we
will see that its overall sample efficiency is much higher than the
earlier methods.

@@

This appendix is divided into two main parts. Appendix~\ref{app:proofs} provides
the proofs that we omitted from the main text due to space constraints.
Appendix~\ref{app:experiments} provides further details on the correction
distribution derivation and on our three main experiments to assist
understanding and reproducibility. 
% Daniel: TODO can decide whether to include this here or not. It might actually
% be better to do this in the logistic regression section.
% Our code is open-source on
% GitHub.\footnote{\url{https://github.com/BIDData/BIDMach/blob/master/src/main/scala/BIDMach/updaters/MHTest.scala}}

@@

>>>>

We note that this approach is fundamentally different from prior
work. It makes no assumptions about the form of, and requires no global bounds on the
posterior parameter distribution. It is exact in the
limit as batch size increases by the Central Limit Theorem. This is
not true of~\citep{cutting_mh_2014} and~\citep{icml2014c1_bardenet14}
which use tail bounds and provide only approximate tests even with
arbitrarily large batches of data. Our test is also exact under the
assumptions of~\citet{cutting_mh_2014} that the log probability ratios
of batches are normally distributed about their mean.  Rather
than tail bounds, our approach uses moment estimates from the data to
determine how far the minibatch posteriors deviate from a normal
distribution.  These bounds carry through to the overall accuracy of
the test.

@@

\subsection{BARKER (LOGISTIC) ACCEPTANCE FUNCTION}\label{ssec:barker_function}
For our new MH test we use the Barker logistic~\citep{Barker65}
function: $g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows
that it satisfies the condition in Lemma~\ref{lem:detailed_balance}.
It is slightly less efficient than the Metropolis test, since its
acceptance rate for vanishing likelihood difference is 0.5. However we
will see that its overall sample efficiency is much higher than the
earlier methods. See Appendix~\ref{app:why_barker} for details.

@@

This appendix is divided into three major sections. Appendix~\ref{app:proofs} provides
the proofs that we omitted from the main text due to space constraints.
Appendix~\ref{app:why_barker} elaborates on our choice of the Barker logistic
function. Finally, Appendix~\ref{app:experiments} presents further details on the correction
distribution numerical derivation and on our three main experiments to assist
understanding and reproducibility.

% Daniel: TODO can decide whether to include this here or not. It might actually
% be better to do this in the logistic regression section.
% Our code is open-source on
% GitHub.\footnote{\url{https://github.com/BIDData/BIDMach/blob/master/src/main/scala/BIDMach/updaters/MHTest.scala}}

@@

\section{WHY THE BARKER LOGISTIC FUNCTION?}\label{app:why_barker}
% Daniel: adding this one due to reviewer concerns

Regarding our choice of the Logistic function, a test function $f(x)$ for
Metropolis-Hastings must satisfy Lemma~\ref{lem:detailed_balance}. In addition,
it must be monotone, bounded by $[0,1]$ and be such that $\lim_{x \to
-\infty}f(x) = 0$ and $\lim_{x \to \infty}f(x) = 1$. While many functions
satisfy this, including the classical test $f(x) = \min\{\exp(x), 1\}$, the
Logistic function is the \emph{unique} function in this class which is
anti-symmetric about 0.5, so it represents the (unique) CDF of a symmetric
random variable. Our method requires approximating this with the sum of a
Gaussian random variable (which is symmetric) and a correction. The Logistic CDF
$L$ and Gaussian CDF $\Phi$ are extremely close even without correction; more
precisely, the CDF error from the closest Gaussian CDF numerically satisfies
$\sup_x |L(x) - \Phi(x/1.7)| < 0.01$, i.e. the sup error between Logistic and
Gaussian CDFs is less than 1\%. With our correction we can make this error
orders of magnitude smaller.

While not a proof of optimality, it is unlikely that a non-symmetric test
function $f(x)$ --- representing a skewed variable --- would do better. It would
require a highly-skewed correction variable, and likely require a much narrower
normal distribution (and hence more samples).
